---
title: "SOC 506 Final Coding"
author: "Amber Burrell"
date: "5/23/2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

## **SOC 506 Final Coding**
## Amber Burrell
## 23 May 2022

```{r setup, include=FALSE}
#install.packages('tidycensus')
library(tidycensus)
library(tidyverse)
#install.packages("haven")
library(haven)
library(MASS)
library(readr)
library(corrplot)
#install.packages("coefplot")
install.packages("ggplot2")
library(ggplot2)

```



This is my 2016 data. below I will be cleaning and manipulating the data to get it ready for a regression.
```{r}
orginial<- read_sav("anes_timeseries_2016.sav")
election2016<- orginial[,c("V161188","V161496", "V167534", "V161310a", "V161310b", "V168307", "V168305")]
election2016<- na.omit(election2016)
election2016$new_black <- ifelse(election2016$V161310b == 1, 1, ifelse(election2016$V161310a == 1, 0, NA))
Ngun<- election2016[,"V161496"]
GunAcc<- election2016[,"V161188"]
neighborhood<- election2016[,"V167534"]
sex <- election2016[,"V168307"]
election2016$GunAcc <- election2016$V161188
election2016$neighborhood <- election2016$V167534
election2016$sex <- election2016$V168307
election2016$Ngun <- election2016$V161496
```

cleaning the data by creating NA for unwanted data and removing it from the dataframe.
```{r}
election2016$GunAcc[election2016$V161188 == -9] <- NA
election2016$GunAcc[election2016$V161188 == -8] <- NA
election2016$Ngun[election2016$V161496 == -9] <- NA
election2016$Ngun[election2016$V161496 == -5] <- NA
election2016$neighborhood[election2016$V167534 == -8] <- NA
election2016$neighborhood[election2016$V167534 == -1] <- NA
election2016$sex[election2016$V168307 == -6] <- NA
election2016$sex[election2016$V168307 == -1] <- NA
clean_2016<- na.omit(election2016)
summary(clean_2016)
```

Many of the variables are categorical or are factor variables so below I am making the variable as factors in R
```{r}
neighborhood_factor <- factor(clean_2016$neighborhood, levels= c(1,2,3,4), labels= c("Farm","Rural","Suburban","Urban"))
sex_factor <- factor(clean_2016$sex, levels= c(1,2), labels= c("Male","Female"))
GunAcc_factor <- factor(clean_2016$GunAcc, levels= c(1,2,3,4,5), labels= c(" Extremely important", "Very important", "Somewhat important", "Not too important", "Not important at all"))

```

The regression and analyzing the regression
```{r}
Guns<- polr(GunAcc_factor~1 + Ngun + new_black + neighborhood_factor, data = clean_2016)
summary(Guns)
exp(coef(Guns))
summary(Guns)$coefficients[,1]
summary(Guns)$standard.error[,2]
z<- summary(Guns)$coefficients[,1]/summary(Guns)$coefficients[,2]
z
pvalue<- (1-pnorm(abs(z),0,1)) *2
pvalue
```

Now here we are making the graphs! 
```{r}
boxplot(GunAcc_factor~Ngun, data = clean_2016 )
boxplot(GunAcc_factor~new_black, data = clean_2016 )
library(sjPlot)
library(coefplot)
plot_model(Guns)
coefplot(Guns)

```



************************************************************



Now we are switching to 2020. similar to 2016, I will be first cleaning and manipulating the data to get ready for the regression.
```{r}
original2020 <- read_csv("anes_timeseries_2020_csv_20220210.csv")
election2020 <- original2020[,c("V202355", "V203412", "V203413","V201628", "V202338" )]
election2020$new_black20 <- ifelse(election2020$V203412 == 2, 1, ifelse(election2020$V203412 == 1, 0, NA))
GunA <- election2020[,"V202338"]
Ngun<- election2020[,"V201628"]
neighborhood<- election2020[,"V202355"]
race <- election2020[,"V203412"]
sex <- election2020[,"V203413"]
election2020$GunA <- election2020$V202338
election2020$neighborhood <- election2020$V202355
election2020$race <- election2020$V203412
election2020$sex <- election2020$V203413
election2020$Ngun <- election2020$V201628
```

Now I will be labeling unwanted data point as NA to then get them cleared to remove all NA from the dataframe to create a clean data frame. 

```{r}
election2020$GunA[election2020$V202338 == -9] <- NA
election2020$GunA[election2020$V202338 == -7] <- NA
election2020$GunA[election2020$V202338 == -6] <- NA
election2020$GunA[election2020$V202338 == -5] <- NA
election2020$Ngun[election2020$V201628 == -9] <- NA
election2020$Ngun[election2020$V201628 == -5] <- NA
election2020$neighborhood[election2020$V202355 == -9] <- NA
election2020$neighborhood[election2020$V202355 == -8] <- NA
election2020$neighborhood[election2020$V202355 == -7] <- NA
election2020$neighborhood[election2020$V202355 == -6] <- NA
election2020$race[election2020$V203412 == -2] <- NA
election2020$race[election2020$V203412 == -1] <- NA
election2020$race[election2020$V203412 == 3] <- NA
election2020$sex[election2020$V203413 == -2] <- NA
election2020$sex[election2020$V203413 == -1] <- NA
clean20<- na.omit(election2020)
```

Now I am creating the factors for the variable for both dependent and independent variable. 
```{r}
neighborhood_f20 <- factor(clean20$neighborhood, levels= c(1,2,3,4), labels= c("Rural","Small Town","Suburb","City"))
sex_f20 <- factor(clean20$sex, levels= c(1,2), labels= c("Male","Female"))
GunA_factor <- factor(clean20$GunA, levels= c(1,2,3,4,5), labels= c(" Extremely important", "Very important", "Somewhat important", "Not too important", "Not important at all"))

```

Regression time! 

```{r}
Guns20<- polr(GunA_factor~1 + Ngun + new_black20 + neighborhood_f20, data = clean20)
summary(Guns20)
exp(coef(Guns20))

z<- summary(Guns20)$coefficients[,1]/summary(Guns20)$coefficients[,2]
z
pvalue<- (1-pnorm(abs(z),0,1)) *2
pvalue

```

Lets make the graphs for the data.

```{r}
boxplot(GunA_factor~Ngun, data = clean20 )
boxplot(GunA_factor~new_black20, data = clean20)
library(sjPlot)
plot_model(Guns20)
library(coefplot)
coefplot(Guns20)

```

Comparison between 2020 and 2016 odds ratio

```{r}
exp(coef(Guns20))
exp(coef(Guns))
(exp(coef(Guns20))) - (exp(coef(Guns)))


```
